FoRTE baseline

### 2019-07-11 analysis/paper/paper.Rmd
Revise first 2 intro paragraphs

**Temperate forests in the Upper Midwest are important and changing quickly.**
{C sequestration}.
{Other ecosystem services}.
However, the fate of these forests is highly uncertain.
The natural ecological succession of these forests has been significantly altered by the interactions of a significantly different and rapidly changing climate, and novel disturbance regimes under which some natural disturbances (e.g. fire) are suppressed while others (e.g. harvest; pests and pathogens) may be becoming more frequent. <!-- TODO refs -->
{UMBS ecological background/natural history?} 
**Therefore, predictions about the future of these forests based on past analogs may not be reliable.**
BBL: This paragraph obviously needs significant fleshing out, but this should be straightforward; I think we can basically take (with a bit of reworking) material from Chris's 2016 _Ecosphere_ paper: http://dx.doi.org/10.1002/ecs2.1375

**More reliable predictions can in theory be obtained by using dynamic vegetation models that explicitly represent processes involved in forest growth and mortality.**
Vegetation models fall broadly along the following spectrum of complexity.
On one side are relatively simple **"Big leaf" models**---such as PNET, SiBCASA, and Biome-BGC <!-- TODO refs; more examples? BBL: no! -->---in which the vegetation at a particular location consists essentially of a single large "plant" whose characteristics are the (weighted) average of all the vegetation at that site. <!-- TODO refs -->
These models were designed originally to simulate the land surface boundary condition for atmospheric general circulation models (BBL: I don't think is is true; earliest models like Forest-BGC were standalone), though similar models continue to be applied successfully to simulate biogeochemical processes in many ecosystem science contexts. <!-- TODO refs -->
On the other end of the spectrum are **"individual-based" models** (a.k.a. "gap models"), which explicitly simulate multiple indivduals competing for resources at a single site [@shugart_2015_computer]. <!-- TODO other refs? -->
Examples of such models are LANDIS <!-- TODO ref !--> and UVAFME. <!-- TODO ref --> <!-- TODO Other examples? -->
Because these models can explicitly represent inter-specific differences in plant productivity, resource allocation, and stress tolerance (among others), as well as the competition that emerges out of these differences, they may be better able to represent changes in ecosystem-scale processes, especially in no-analog conditions [@purves_2008_predictive].
In the middle are models that capture the emergent biogeochemical and ecological outcomes of interactions between individual plants without the need for explicitly simulating each individual [@purves_2008_predictive], and have less-intensive data requirements generally.
In this study, we focus on the Ecosystem Demography model [ED2.2; @moorcroft_2001_method; @medvigy_2011_predicting; @longo_2019_ed1], which constitutes an example of such an approach.

>>>>

Forests are a defining feature of the US North.
This region is the most heavily forested region in the US, and these forests are an essential provider of employment and recreation opportunities and other ecosystem services for the region's people, and of habitat for the region's wildlife [@shividenko_2005_forest; @shifley_2012_forests].
In addition, the forests of the Upper Midwest in particular have been a net carbon sink for most of the past century, as fast-growing early successional species established and thrived following near-complete deforestation around the end of the 19th Century [@birdsey_2006_forest; @williams_2012_carbon].
However, the fate of these forests is highly uncertain.
For one, many Upper Midwest forests are undergoing a shift in community composition from early-successional to mid-successional species, and the ecological and biogeochemical consequences of this transition are not well-established [@gough_2016_disturbance].
In addition, these forests face a multitude of direct and indirect pressures from both biotic and abiotic sources [@shifley_2016_future], including longer and more severe droughts [@gustafson_2016_implications; @swanston_2017_vulnerability; @lienard_2016_us], non-native insects and pathogens [@lovett_2016_nonnative], and more. <!-- TODO More here? -->
Collectively, these processes have culminated in forests that are different from their pre-settlement climax counterparts [@stearns_2002_one; @wolter_2002_recent; @thompson_2013_four], and our ability to make predictions about such non-analog environments based on the past is limited.

**More reliable predictions can in theory be obtained by using dynamic vegetation models that explicitly represent processes involved in forest growth and mortality.**
Vegetation models fall broadly along the following spectrum of complexity.
On one side are relatively simple **"big leaf" models**---such as PNET, SiBCASA, and Biome-BGC---in which the vegetation at a particular location consists essentially of a single large "plant" whose characteristics are the (weighted) average of all the vegetation at that site.
A common application of these models was simulating the land surface boundary condition for atmospheric general circulation models, though similar models have been (and continue to be) applied successfully to simulate biogeochemical processes in many ecosystem science contexts. <!-- TODO refs -->
On the other end of the spectrum are **"individual-based" models** (a.k.a. "gap models"), which explicitly simulate multiple indivduals competing for resources at a single site [@shugart_2015_computer]. <!-- TODO other refs? -->
Examples of such models are LANDIS [@scheller_2007_design] and UVAFME. <!-- TODO ref --> <!-- TODO Other examples? -->
Because these models can explicitly represent inter-specific differences in plant productivity, resource allocation, and stress tolerance (among others), as well as the competition that emerges out of these differences, they may be better able to represent changes in ecosystem-scale processes, especially in no-analog conditions [@purves_2008_predictive]

In the middle are models that capture the emergent biogeochemical and ecological outcomes of interactions between individual plants without the need for explicitly simulating each individual [@purves_2008_predictive], and have less-intensive data requirements generally.
In this study, we focus on the Ecosystem Demography model [ED2.2; @moorcroft_2001_method; @medvigy_2011_predicting; @longo_2019_ed1], which constitutes an example of such an approach.
